---
output: github_document
---

```{r include = FALSE}
library("targets")
```


# Translating priors from linear mixed models to repeated-measures ANOVA and paired *t* tests

**Frederik Aust, Johnny van Doorn & Julia M. Haaf**

This repository contains research products associated with the publication.
We provide a Dockerfile, R scripts, and a Quarto document to reproduce the reported simulations.
The Quarto document in the `reports` directory and the `_targets_*.R` scripts contain details about reported simulations and can be used to reproduce the results.
With the help of [Quarto](https://quarto.org/) the `.qmd`-file can be rendered into a report of the simulation results in HTML format.

View the report of the simulation results [here](https://htmlpreview.github.io/?https://github.com/crsh/bf-defprit/blob/main/report/prior_translation.html).


## Recommended citation

Aust, van Doorn, & Haaf (2022). Translating priors from linear mixed models to repeated-measures ANOVA and paired *t* tests. *PsyArXiv*.

## Software requirements

The required software is detailed in the `DESCRIPTION` file.

To run the Docker container, execute `_run_docker.sh`, e.g. `sh _run_docker.sh` in the terminal.
This will recreate the software environment used to run the simulation.
The environment can be interacted with in an RStudio instance in a web browser that will open automatically.

## `targets` pipelines

The simulations were run using reproducible pipelines defined with the `targets` package.
To run the pipeline execute `_make.sh`, e.g. `sh _make.sh` in the terminal.
This project contains three separate but interdependent pipelines

### Partial aggregation simulation

```{r echo = FALSE, results = "asis"}
cat("```mermaid\n")
targets::tar_mermaid(
  script = "_targets_partial_aggregation.R"
  , store = "_targets_partial_aggregation"
  , reporter = "silent"
  ) |> 
  cat(sep = "\n")
cat("\n```")
```

### Full aggregation simulation

```{r echo = FALSE, results = "asis"}
cat("```mermaid\n")
targets::tar_mermaid(
  script = "_targets_full_aggregation.R"
  , store = "_targets_full_aggregation"
  , reporter = "silent"
  ) |> 
  cat(sep = "\n")
cat("\n```")
```

### Report

```{r echo = FALSE, results = "asis"}
cat("```mermaid\n")
targets::tar_mermaid(
  script = "_targets_report.R"
  , store = "_targets_report"
  ) |> 
  cat(sep = "\n")
cat("\n```")
```

## Simulation data

The `_targets*/objects` directories contain all (intermediary) simulation results in `RDS`-format, which can be readily imported into R using `targets::tar_load()` or `loadRDS()`.
